An Empirical Analysis of the Internet Engineering Task Force with Computational Methods

### 2024-10-03 paper.tex
Tweak intro + abstract

  The Internet requires interoperability between networks, systems, and
  applications, as well as cooperation among a growing number of
  stakeholders. The Internet Engineering Task Force (IETF) is critical in
  supporting this cooperation and interoperability by bringing together
  interested parties and standardising the protocols that power the
  Internet, such as IPv4/6 or HTTP/s. Our work covers more 20 years of IETF
  data, and analyses the participants of the IETF, the emails they
  exchange, and the documents they produce. We show how these data can be
  used to understand who are these Internet actors, and their interests as
  well as how Internet's growth and maturity has given rise to a longer and
  complex standardisation process where participants gain and exercise
  influence and how this is reflected in the language they use and the
  interactions they have.

@@

% A good paper introduction is fairly formulaic. If you follow a simple set
% of rules, you can write a very good introduction. The following outline can
% be varied. For example, you can use two paragraphs instead of one, or you
% can place more emphasis on one aspect of the intro than another. But in all
% cases, all of the points below need to be covered in an introduction, and
% in most papers, you don't need to cover anything more in an introduction.
%
% Paragraph 1: Motivation. At a high level, what is the problem area you
% are working in and why is it important? It is important to set the larger
% context here. Why is the problem of interest and importance to the larger
% community?

@@

>>>>

  The Internet requires interoperability between networks, systems, and
  applications, as well as cooperation among a growing number of
  stakeholders. The Internet Engineering Task Force (IETF) is critical in
  supporting this cooperation and interoperability by bringing together
  interested parties and standardising the protocols that power the
  Internet, such as IP or HTTP. Our work covers more 20 years of IETF
  data, and analyses the participants of the IETF, the emails they
  exchange, and the documents they produce. We show how these data can be
  used to understand who are these Internet actors, and their interests as
  well as how Internet's growth and maturity has given rise to a longer and
  complex standardisation process where participants gain and exercise
  influence and how this is reflected in the language they use and the
  interactions they have.

@@

% Paragraph 1: Motivation. At a high level, what is the problem area you
% are working in and why is it important? It is important to set the larger
% context here. Why is the problem of interest and importance to the larger
% community?

Network protocol standards are crucial for successful operation of the
Internet.  A successful standard provides a basis for interoperability
between systems developed by competing vendors, and supports the growth of
an open ecosystem of products and services. Further, the process by which
network protocol standards are developed, comprising multiple rounds of
open feedback and review, has proven remarkably effective in designing
high-quality and robust protocols, many of which see widespread deployment
and use. Understanding the Internet standards development process, and how
it produces successful protocols is, therefore, important if we are to
understand the Internet and how it has evolved.

One of the main organisations that develops protocol standards is the
Internet Engineering Task Force (IETF). The IETF was founded in 1986,
following on from the US Government-funded effort that developed the early
Internet. It has since grown to become a global community of network
protocol designers, vendors, network operators, and researchers that
develop and publish open network protocol standards and operational
guidelines.  The IETF publishes its standards, and other documents, in
the RFC series.\footnote{\url{https://www.rfc-editor.org}} This series
comprises around 9,000 documents, authored over more than 50 years,
and provides a rich history of the development of the Internet and
its protocols \cite{RFC8700}.

@@

Standardisation is an inherently social and political process
\cite{RFC2026,RFC7282}, requiring cooperation and consensus among a growing
number of stakeholders.

It is important to gain a coherent understanding of the activities that
take place within the IETF, as well as the key factors that may predict
the success of a protocol standard.
